---
title: "Experiment 1 - Dice, Deck of Cards, Slot Machine"
author: "Saha Debanshee Gopal"
date: "26th August 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**A) DICE**

**1) Problem Statement: To plot x and y using qplot**

**Code:**
```{r}
x <- c(-1, -0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8, 1)
x
y <- x^3
y
```
*Here, using th basic computation of R, the cube of x is stored as the value of y.*

**Output:**
```{r}
library(ggplot2)
qplot(x, y)
```


*A scatterplot is a set of points, each plotted according to its x and y values. Together, the vectors x and y describe a set of 10 points. Scatterplots are useful for visualizing the relationship between two variables.*

**2) Problem Statement: To plot the probability of a normal dice:**

**Code:**
```{r}
roll<-function(){
die<-1:6
dice<-sample(die,size=2,replace=TRUE)
sum(dice)
}
rolls<-replicate(10000,roll())
```

**Output:**
```{r}
library(ggplot2)
qplot(rolls,binwidth=1)
```

*A histogram visualizes the distribution of a single variable. It displays how many data points appear at each value of x. The behavior of our dice suggests that they are fair. Seven occurs more often than any other number, and frequencies diminish in proportion to the number of die combinations that create each number*

**3) Problem Statement: To plot the probability of a weighted dice:**

**Code:** 
```{r}
roll1<-function(){
die<-1:6
dice<-sample(die,size=2,replace=TRUE,
prob = c(1/8, 1/8, 1/8, 1/8, 1/8, 3/8))
sum(dice)
}
rolls1<-replicate(10000,roll1())
```
**Output:**
```{r}
library(ggplot2)
qplot(rolls1,binwidth=1)
```

*The dice are now clearly biased towards high numbers, since high sums occur much more often than low sums*


#**B) DECK OF CARDS**

**1) Problem Statement: Define suits, cards, values**

**Code**
```{r}
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
face <- c("Ace", "Deuce", "Three", "Four","Five", "Six", "Seven", "Eight",
           "Nine", "Ten", "Jack", "Queen", "King")
values <- c(rep(1:13,4))
```

**2) Problem Statement: To show the deck of cards along with its value using data.frame**

**Code:**
```{r}
deck <- data.frame(
face = c("king", "queen", "jack", "ten", "nine", "eight", "seven", "six",
"five", "four", "three", "two", "ace", "king", "queen", "jack", "ten",
"nine", "eight", "seven", "six", "five", "four", "three", "two", "ace",
"king", "queen", "jack", "ten", "nine", "eight", "seven", "six", "five",
"four", "three", "two", "ace", "king", "queen", "jack", "ten", "nine",
"eight", "seven", "six", "five", "four", "three", "two", "ace"),
suit = c("spades", "spades", "spades", "spades", "spades", "spades",
"spades", "spades", "spades", "spades", "spades", "spades", "spades",
"clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs", "clubs",
"clubs", "clubs", "clubs", "clubs", "clubs", "diamonds", "diamonds",
"diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "diamonds",
"diamonds", "diamonds", "diamonds", "diamonds", "diamonds", "hearts",
"hearts", "hearts", "hearts", "hearts", "hearts", "hearts", "hearts",
"hearts", "hearts", "hearts", "hearts", "hearts"),
value = c(13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8,
7, 6, 5, 4, 3, 2, 1, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1, 13, 12, 11,
10, 9, 8, 7, 6, 5, 4, 3, 2, 1))
```
**Output:**
```{r}
deck
```
**3) Problem Statement: To shuffle and deal a deck**

**Code:**
```{r}
deal <- function(cards) {
cards[1, ]
}
shuffle <- function(cards) {
random <- sample(1:52, size = 52)
cards[random, ]
}
```
**Output:**
```{r}
deal(deck)
deck2 <- shuffle(deck)
```
**Output 2:**
```{r}
deal(deck2)
```

**4) Problem Statement: Creating Deck**

**Code:**
```{r}
suits <- c("Diamonds", "Clubs", "Hearts", "Spades")
face <- c("Ace", "Deuce", "Three", "Four","Five", "Six", "Seven", "Eight",
          "Nine", "Ten", "Jack", "Queen", "King")
values <- c(rep(1:13,4))

deck <- expand.grid(face=face, suits=suits)
deck$value <- values


deck2 <- deck

deck2$new <- 1:52

deck2[c(13, 26, 39, 52), ]
```

**5) Problem Statement: Count the number of Ace in Deck**

**Code:**
```{r}
deck2$face == "Ace"
sum(deck2$face == "Ace")
```

**6) Problem Statement: Assign value 14 to every Ace in deck3**

**Code**
```{r}
deck3 <- deck
deck3$value[deck3$face == "Ace"] <- 14

```

**7) Problem Statement: Creating Deck for Hearts**

**Code**
```{r}
deck4<-deck

deck4$value <- 0
deck4$value[deck4$suits == "Hearts"] <- 1
deck4$value[deck4$suits == "Hearts"]

deck4$face == "Queen" & deck4$suits == "Spades"

queenofsp <- deck4$face == "Queen" & deck4$suits == "Spades"
deck4[queenofsp, ]

deck4$value[queenofsp] <- 13 
deck4$face[queenofsp] <- "Queen"
deck4$suits[queenofsp] <- "Spades"
```
**Output:**
```{r, results='asis'}
knitr::kable(deck4)
```

**8) Problem Statement: Keeping original deck safe (Closure)**

**Code:**
```{r}
setup <- function(deck) {
DECK <- deck
DEAL <- function() {
card <- deck[1, ]
assign("deck", deck[-1, ], envir = parent.env(environment()))
card
}
SHUFFLE <- function(){
random <- sample(1:52, size = 52)
assign("deck", DECK[random, ], envir = parent.env(environment()))
}
list(deal = DEAL, shuffle = SHUFFLE)
}
cards <- setup(deck)
deal <- cards$deal
shuffle <- cards$shuffle
```
*Closure ensures that even if we remove the original deck, we can continue playing cards.*

\newpage

#**C) SLOT MACHINE**

*A code in R which allows us to play the most popular mordern casino game.* 

*Symbols used include the following:*

*DD  - Diamonds    (0.03)*

*7   - Seven       (0.03)*

*BBB - Triple Bars (0.06)*

*BB  - Double Bars (0.1)*

*B   - Single Bars (0.25)*

*C   - Cherries    (0.01)*

*0   - Zeros       (0.52)*

*with the probabilities in the brackets.*


**1) Problem Statement: To select symbols randomly using the sample function**

**Code**
```{r}
get_symbols <- function() {
  wheel <- c("DD","7","BBB","BB","B","C","0")
  sample(wheel,size = 3,replace= TRUE,prob = c(0.03,0.03,0.06,0.1,0.25,0.01,0.52))
}
```
**Output:**
```{r}
get_symbols()
```
*For the actual game, we need to assign score to the symbols and that can be done via the score function. The score of the 3 random symbols obtained from 'get_symbols' is extracted from a lookup table which has all the values.*


**Code:** 
```{r}
score <- function(symbols){
  same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
  bars <- symbols %in% c("B","BB","BBB")
  if(same){
    payouts <- c("DD" = 100,
                 "7" = 80,
                 "BBB" = 40,
                 "BB" = 25,
                 "B" = 10,
                 "C" = 10,
                 "0" = 0)
    prize <- unname(payouts[symbols[1]])
  }
  else if(all(bars)){
    prize <- 5
  }
  else{
    cherries <- sum(symbols == "C")
    prize <- c(0,2,5)[cherries +1]
  }
  
  diamonds <- sum(symbols == "DD")
  prize * 2 ^ diamonds
}
```
*The game can be run using this function which calls the get_symbols() functions and the score() function.*

**Code:**
```{r}
play <- function(){
  symbols <- get_symbols()
  print(symbols)
  score(symbols)
}
play()
```
*Modified play() function to store the values of the symbols as attributes with the value of prize.*

**Code:**
```{r}
play <- function(){
  symbols <- get_symbols()
  structure(score(symbols),symbols = symbols)
}
play()
```
*The structure function creates an object with a set of attributes. The first argument should be a R object or set of values and the remaining arguments should be named attributes for the structure to add to the object.*

*The attributes now can be used to create a slot_display() functiona as follows:*

```{r}
slot_display <- function(prize){
  #extract the symbols
  symbols <- attr(prize,"symbols")    
  #combine symbol with prize as a regular expression
  symbols <- paste(symbols,collapse = " ")
  #append with new line character
  string <- paste(symbols,prize,sep="\n$")  
  cat(string)   #display without quotes
}
one_play <- play()
```

*We use expand.grid to find out all the possible combinations of a vector with another vector. Using this we calculate the possible combinations of the wheel.*

```{r}
wheel <- c("DD","7","BBB","BB","B","C","0")
combos <- expand.grid(wheel,wheel,wheel,stringsAsFactors = FALSE)
head(combos,3)
```
*This creates a variable combos with 343 observations.*

*We then create a new lookup table for the probabilites and add those values to combos as a factor and, calculate and add a total probability for each of the combination.*

```{r}
prob <- c("DD" = 0.03,"7" = 0.03,"BBB" = 0.06,"BB" = 0.1,"B" = 0.25,"C" = 0.01,"0" = 0.52)

combos$prob1 <- prob[combos$Var1]
combos$prob2 <- prob[combos$Var2]
combos$prob3 <- prob[combos$Var3]

combos$prob = combos$prob1 * combos$prob2 * combos$prob3
head(combos,3)
```
*To store the prize along with the combinations, we use a for loop and add prize as a factor.*

```{r}
combos$prize <- NA
for(i in 1:nrow(combos)){
  symbols <- c(combos[i,1],combos[i,2],combos[i,3])
  combos$prize[i] <- score(symbols)
}
head(combos,3)
```
*The expected value of the prize won is given by*

```{r}
sum(combos$prize * combos$prob)
```
*This value is less than the value mentioned by the manufacturer because of the wild card "DD".*

*Correcting the code to fix the wild card problem and recalculating the expected value of prize.*

```{r}
score <- function(symbols){
  diamonds <- sum(symbols == "DD")
  cherries <- sum(symbols == "C")
  #case identification
  slots <- symbols[symbols != "DD"]
  same <- length(unique(slots)) == 1
  bars <- slots %in% c("B","BB","BBB")
  #assign prize
  if(diamonds == 3){
    prize <- 100
  } else if(same){
    payouts <- c("7"=80,"BBB"=40,"BB"=25,"B"=10,"C"=10,"0"=0)
    prize <- unname(payouts[slots[1]])
  } else if(all(bars)){
    prize <- 5
  } else if(cherries > 0){
    prize <- c(0,2,5)[cherries + diamonds + 1]
  } else {
    prize <- 0
  }
  #double the prize for each diamond
  prize * 2^diamonds  
}

#reassinging the prize value
combos$prize <- NA
for(i in 1:nrow(combos)){
  symbols <- c(combos[i,1],combos[i,2],combos[i,3])
  combos$prize[i] <- score(symbols)
}

#finding sum
sum(combos$prize * combos$prob)
```
**2) Problem Statement: To demonstrate the playing in till the cash runs out** 

**Code:**
```{r}
plays_till_broke <- function(start_with) {
  cash <- start_with
  n<-0
  while(cash >0){
    cash <- cash - 1 + play()
    n <- n + 1
  }
  n
}
```


```{r fig.width=2,fig.height=5,echo=FALSE}
library(ggplot2)
```